MA_Results

### 2022-05-31 Markdown/Manuscript/Child_RMDs/MS_Discussion.Rmd
circulated draft

MH systems design is not a pharma led project - less concerns about commercial ownership

greater use of these types of models may require adaptation on the part of funders, modellers and decision-makers. T

>>>>

<!-- MH systems design is not a pharma led project - less concerns about commercial ownership -->

<!-- greater use of these types of models may require adaptation on the part of funders, modellers and decision-makers. T -->

### 2022-05-31 Markdown/Manuscript/Child_RMDs/MS_Intro_Short.Rmd
circulated draft

<!-- `r ifelse(is.na(params$X@background_1L_chr),"",params$X@background_1L_chr)`  -->
Health economic computational models have become essential tools for health policy development [@dakin2015influence; @Erdemir2020]. Although influential and widely used, these models routinely contain errors [@Radeva2020], are rarely adequately validated [@Ghabri2019], can be difficult to reproduce [@Jalali2021; @McManus2019; @Bermejo2017] and are likely to be infrequently updated or revised [@Sampson2017]. To help address these issues, there is growing support for greater use of open source health economic models (OSHEMs) that grant liberal permissions to access and re-use model source code [@Pouwels2022]. However, to date actual implementations of OSHEMs are rare [@Emerson2019; @Michalczyk2018; @Feenstra2022]. Barriers to adoption include concerns about intellectual property, confidentiality, model misuse and the resources required to support open source implementations [@Pouwels2022; @Wu2019]. As many health economic models are owned by pharmaceutical companies and consultancies, commercial considerations may also limit the uptake of OSHEMs [@Feenstra2022].

>>>>

<!-- `r ifelse(is.na(params$X@background_1L_chr),"",params$X@background_1L_chr)`  -->
Health economic computational models have become essential tools for health policy development [@dakin2015influence; @Erdemir2020]. Although influential and widely used, these models routinely contain errors [@Radeva2020], are rarely adequately validated [@Ghabri2019], can be difficult to reproduce [@Jalali2021; @McManus2019; @Bermejo2017] and are likely to be infrequently updated or revised [@Sampson2017]. To help address these issues, there is growing support for greater use of open source health economic models (OSHEMs) that grant liberal permissions to access and re-use model source code and data [@Pouwels2022]. However, to date actual implementations of OSHEMs are rare [@Emerson2019; @Michalczyk2018; @Feenstra2022]. Barriers to adoption include concerns about intellectual property, confidentiality, model misuse and the resources required to support open source implementations [@Pouwels2022; @Wu2019]. As many health economic models are owned by pharmaceutical companies and consultancies, commercial considerations may also limit the uptake of OSHEMs [@Feenstra2022].

### 2022-05-31 Markdown/Manuscript/Child_RMDs/MS_Results.Rmd
circulated draft

```{r child="../Child_RMDs/MS_Tables.Rmd", echo=FALSE}
```
```{r child="../Child_RMDs/MS_Figures.Rmd", echo=FALSE}
```
<!-- Add child RMDs for results content -->

@@

The development, validation and updating of more complex mental health economic models implemented over longer time frames may be too onerous a burden for a single modelling team. Developing networks of modellers working on common health conditions has been recommended as a strategy for improving model validity [@Sampson2019] and a some of us are part of a nascent initiate of this type in mental health [@nmhsmn2022]. Furthermore, collaborations across multiple modelling teams that include the ability to re-use and extend each others work can make complex modelling projects more tractable [@Arnold2010]. <!-- Add ref to NMHSMN in Australia --> Similarly, developing partnerships between modellers and decision-makers across the life-cycle of a modelling project can help ensure models are appropriately conceptualised and implemented and improve their practical utility as decision aids [@Zabell2021; @SQUIRES2016588].

@@

Our approach to model development is to undertake a number of discrete modelling projects of the people, places, platforms and programs that shape the mental health and wellbeing of young people and to progressively link them together by means of a common framework. To model people we are developing synthetic representations of populations of interest [@DVN/HJXYKQ_2021] that describe relevant individual characteristics and their household relationships, algorithms that map psychological measures to health utility [@Hamilton2021.07.07.21260129] and choice models for predicting the helpseeking behaviour of young people. Our in development model of places [@DVN/V3OKZV_2022] aims to synthesise geometry and spatial attribute data to characterise the geographic distribution of relevant demographic, environmental, epidemiological and service infrastructure features and our first model of a service platform will represent the processes and operations of a complex primary youth mental health service. We also plan to extend our prior work reviewing economic evidence relating to youth mental health programs [@RN33] so that it can be integrated with the model.

@@

## Standards
We have identified 20 standards that we believe are important for quality implementations of OSHEMs, each described under one of following six principles for making models TIMELY:

@@

Model code and data also need to be clearly documented, potentially with different versions for technical and non-technical users [@Eddy2012]. Developers storing data in a Dataverse installation have access to multiple meta-data fields to document both a data collection and its constituent individual files. In R, code manuals and websites can be created with the aid of tools such as devtools [@devtools2021], sinew [@sinew2022], roxygen2 [@roxygen2021] and pkgdown [@pkgdown2022].

@@

### Licensed Models
```{r}
# Adapted fom: https://seankross.com/2016/08/02/How-R-Packages-are-Licensed.html
sorting_hat <- function(student, houses){
  choice <- purrr::map_lgl(houses, grepl, x = student)
  if(!any(choice)){
    return("Other")
  } else {
    return((houses[choice])[1])
  }
}
license_table <- table(utils::available.packages()[,"License"])
names(license_table) %<>%
  purrr::map_chr(function(x){
    sorting_hat(x, c("Apache", "Artistic", "CC", "BSD", "MIT", "GPL"))
  })  
license_tbl <- as.data.frame(license_table) %>%
  dplyr::rename(License = Var1) %>%
  dplyr::group_by(License) %>%
  dplyr::summarize(Freq = sum(Freq)) %>%
  dplyr::ungroup() %>%
  dplyr::arrange(dplyr::desc(Freq))
#licenses <- license_tbl$Freq
gpl_pc_dbl <- round(license_tbl$Freq[license_tbl$License=="GPL"]/sum(license_tbl$Freq)*100,2) 
```

@@

For code, it may be appropriate to adopt the prevailing open source licensing practice within the programming language being used. Applying a previously published algorithm [@kross2016] to analyse the most comprehensive archive of released R packages [@CRAN2022] finds that `r `gpl_pc_dbl`% are distributed under various forms of General Public License (GPL) [@GNUGPL2022], a copyleft license.

@@

Licenses may or may not specify that model re-users must give appropriate acknowledgement to model authors. Citation tools can be distributed with each cite individual code or data items to inform re-users of the desired attribution. In R, including a CITATION file in the `inst` directory of a package will enable users of that package to retrieve citation information by running a command of the format `citation("Package Name")` in the R console [@Salmon2021]. More generally, including a CITATION.cff file at the top level of your code repository will enable GitHub and Zenodo repositories hosting that item to include the relevant information in their citation tools [@Druskat_Citation_File_Format_2021]. Datasets hosted on Dataverse installations have metadata fields that, once completed by authors, generate citation files for dataset viewers.

@@

## Modelling tools
We have developed a toolkit to help streamline the process of developing OSHEMs that meet the TIMELY standards. The toolkit is comprised of online repositories and software.

@@

- self-documenting model modules [@ready4class2022];

- self-documenting functions (including methods), written in a consistent house style [@ready4fun2022];

@@

The six R packages, their primary focus, the TIMELY standards they support and the third party packages they depend on are summarised in Table \@ref(tab:cpkgs). When used in conjunction with framework repositories, the six packages currently provide strong support for implementing 17 of the TIMELY standards, weak support for implementing the standards relating to disseminating statistical models (Y1) and user-interface development (Y2) and no support for implementing standard T5 (reporting code coverage). Our future development priorities for these packages include adding functionality so that all 20 TIMELY standards are well supported and to then submit production versions of each package for review by and archiving on the Comprehensive R Archive Network (CRAN) [@CRAN2022].

@@

# Application

Worked example

```{r tab.cap='TIMELY Checklist with utility mapping study worked example', tab.id = 'checklisttb', echo = F, results = 'asis'}
    #  a <- c("TEST1","TEST2")
    # b <- c("\\cite{ready4show2022}", "hjhjhj")
    # tmp <- data.frame(a,b)   
#     kableExtra::kbl(tmp, booktabs = T, escape = F) %>%
#  kableExtra::kable_styling(
#    latex_options = c("striped", "hold_position", "repeat_header")
#    ) %>%  
#   kableExtra::column_spec(2, width = "50em") %>%
# kableExtra::landscape()

checklist_df <- data.frame(Standard = c(paste0("T",1:6, c(" (Publicly archived)",
                                                          " (Documented)",
                                                          " (Simple Syntax)",
                                                          " (Literate programming)",
                                                          " (Code coverage)",
                                                          " (Reproducibility")),
                                        paste0("I",1:4, c(" (Version controlled)",
                                                          " (Semantic versioning)",
                                                          " (Continuous integration)",
                                                          " (Deprecation)")),
                                        paste0("M",1:2, c(" (Decouple code and data)",
                                                          " (Encapsulation)")),
                                        paste0("E",1:3, c(" (Code libraries)",
                                                          " (Inheritance)",
                                                          " (Test data)")),
                                        paste0("L",1:3, c(" (Copyleft for code)",
                                                          " (Liberal, safe data terms)",
                                                          " (Citation tools)")),
                                        paste0("Y",1:2, c(" (Statistical model tools)",
                                                          " (User-interfaces"))),
                           Assessment = rep("Met so well it hurts",20),
                           Description = c("5 libraries [@hamilton_matthew_2022_6084467;@hamilton_matthew_2022_6084824;@hamilton_matthew_2022_6116701;@gao_caroline_2022_6130155;@matthew_p_hamilton_2021_5646669], 3 programs [@hamilton_matthew_2022_6129906;@hamilton_matthew_2022_6416330;@hamilton_matthew_p_2022_6321821], 2 sub-routines [@hamilton_matthew_2022_6116385;@matthew_p_hamilton_2022_5976988] and 2 datasets [@DVN/DKDIB0_2021;@DVN/HJXYKQ_2021]",
# @hamilton_matthew_2022_6084467;@hamilton_matthew_2022_6084824;@hamilton_matthew_2022_6116701;@gao_caroline_2022_6130155;@matthew_p_hamilton_2021_5646669
# Programs:

# 
# Sub-routines:

                                           "XX",
                                           "XX",
                                           "XX",
                                           "XX",
                                           "XX",
                                           "XX",
                                           "XX",
                                           "XX",
                                           "XX",
                                           "XX",
                                           "XX",
                                           "XX",
                                           "XX",
                                           "XX",
                                           "XX",
                                           "XX",
                                           "XX",
                                           "XX",
                                           "XX"))

tb2_nms_chr<-colnames(checklist_df)
checklist_df <- data.frame(checklist_df[,1],
                 rep("&nbsp; &nbsp; &nbsp;",nrow(checklist_df)),
                 checklist_df[,2],
                 rep("&nbsp; &nbsp; &nbsp;",nrow(checklist_df)),
                 checklist_df[,3])
colnames(checklist_df) <- c("","Standard",
                            "Met?", " ",
                            tb2_nms_chr[3])
if(params$X@outp_formats_chr[1] == "Word") {
  checklist_df %>%
    ready4show::print_table(output_type_1L_chr = params$X@outp_formats_chr[1],
                            caption_1L_chr = knitr::opts_current$get("tab.cap"),
                            mkdn_tbl_ref_1L_chr = paste0("tab:",knitr::opts_current$get("tab.id")),
                            use_rdocx_1L_lgl = ifelse(params$X@outp_formats_chr[1]=="Word",T,F),
                            add_to_row_ls = NULL,
                            sanitize_fn = NULL)
  }else{
    checklist_df %>%
    kableExtra::kbl(format = "markdown",
                      #escape = F,
                    caption = knitr::opts_current$get("tab.cap"))
    # df %>%
    #   kableExtra::kbl(booktabs = T,
    #                   caption = knitr::opts_current$get("tab.cap")) %>%
    #   kableExtra::kable_styling(full_width = F) %>%
    #   kableExtra::column_spec(9, width = "17em") %>%
    #   kableExtra::add_header_above(c(" " = 2, "Supports standard" = 6, " "))
    }
```

>>>>

```{r child="../Child_RMDs/XX_Funs.Rmd", echo=FALSE}
```
```{r child="../Child_RMDs/MS_Tables.Rmd", echo=FALSE}
```
```{r child="../Child_RMDs/MS_Figures.Rmd", echo=FALSE}
```
<!-- Add child RMDs for results content -->

@@

The development, validation and updating of more complex mental health economic models implemented over longer time frames may be too onerous a burden for a single modelling team. Developing networks of modellers working on common health conditions has been recommended as a strategy for improving model validity [@Sampson2019] and some of us are part of a nascent initiate of this type in mental health [@nmhsmn2022]. Furthermore, collaborations across multiple modelling teams that include the ability to re-use and extend each others work can make complex modelling projects more tractable [@Arnold2010]. <!-- Add ref to NMHSMN in Australia --> Similarly, developing partnerships between modellers and decision-makers across the life-cycle of a modelling project can help ensure models are appropriately conceptualised and implemented and improve their practical utility as decision aids [@Zabell2021; @SQUIRES2016588].

@@

Our approach to model development is to undertake a number of discrete modelling projects of the people, places, platforms and programs that shape the mental health and wellbeing of young people and to progressively link them together by means of a common framework. To model people we are developing synthetic representations of populations of interest [@DVN/HJXYKQ_2021] that describe relevant individual characteristics and their relationships, algorithms that map psychological measures to health utility [@Hamilton2021.07.07.21260129] and choice models for predicting the helpseeking behaviour of young people. Our in development model of places [@DVN/V3OKZV_2022] aims to synthesise geometry and spatial attribute data to characterise the geographic distribution of relevant demographic, environmental, epidemiological and service infrastructure features. We are in the early stages of a multi-annual project to develop a service platform model that will represent the processes and operations of a complex primary youth mental health service. We also plan to extend and update our prior work reviewing economic evidence relating to youth mental health programs [@RN33] so that it can be integrated with the model.

@@

## Framework standards
We have identified 20 standards that we believe are important for quality implementations of OSHEMs, each described under one of following six principles for making models TIMELY:

@@

Model code and data also need to be clearly documented, potentially with different versions for technical and non-technical users [@Eddy2012]. Developers storing data in a Dataverse installation have access to multiple meta-data fields to document both a data collection and its individual constituent files. In R, code manuals and websites can be created with the aid of tools such as devtools [@devtools2021], sinew [@sinew2022], roxygen2 [@roxygen2021] and pkgdown [@pkgdown2022].

@@

### Licensed Models
```{r echo = F}
gpl_pc_dbl <- get_license_share("GPL")
```

@@

For code, it may be appropriate to adopt the prevailing open source licensing practice within the programming language being used. Applying a previously published algorithm [@kross2016] to analyse the most comprehensive archive of released R packages [@CRAN2022] finds that `r gpl_pc_dbl`% are distributed under various forms of General Public License (GPL) [@GNUGPL2022], a copyleft license.

@@

Licenses may or may not specify that model re-users must give appropriate acknowledgement to model authors. Citation tools can be distributed with each individual code or data item to inform re-users of the desired attribution. In R, including a CITATION file in the `inst` directory of a package will enable users of that package to retrieve citation information by running a command of the format `citation("Package Name")` in the R console [@Salmon2021]. More generally, including a CITATION.cff file at the top level of your code repository will enable GitHub and Zenodo repositories hosting that item to include the relevant information in their citation tools [@Druskat_Citation_File_Format_2021]. Datasets hosted on Dataverse installations have metadata fields that, once completed by authors, generate citation files for dataset viewers.

@@

## Framework toolkits
We have developed a toolkit to help streamline the process of developing OSHEMs that meet the TIMELY standards. The toolkit is comprised of online repositories and software.

@@

- documented model modules [@ready4class2022];

- documented functions (including methods), written in a consistent house style [@ready4fun2022];

@@

The six R packages, their primary focus, the TIMELY standards they support and the third party packages they depend on are summarised in Table \@ref(tab:cpkgs). When used in conjunction with framework repositories, the six packages extending existing R packages to provide strong support for implementing 16 of the TIMELY standards. However, the software only weakly supports implementing the standards relating to disseminating statistical models (Y1) and user-interface development (Y2) and does not yet provide any workflow tools to help implement the standards for code coverage (T5) and deprecation conventions (I4). Standards not supported or weakly supported by our software can be met with existing developer tools in R and we plan to progressively integrate these third party tools with our own in future releases of our software. Another future priority is to submit production versions of each R package for review by and archiving on the Comprehensive R Archive Network (CRAN) [@CRAN2022].

@@

\blandscape

```{r, child = child_docs_ls$two, echo=FALSE}
```

\elandscape

# Application
We used framework toolkits to develop model code and datasets for implementing a utility mapping study [@Hamilton2021.07.07.21260129]. Table \@ref(tab:checktb) assesses that study against each TIMELY standard, 18 of which the study was able to meet. The two current standards where the study falls short are in reporting code coverage and including a user-interface. Both these items are current development priorities for our team.
